04_LDA_v2

### 2019-12-23 analysis/04_LDA_v2.Rmd
new lda

---
title: "04_DESeq_sexes"
output: md_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(cowplot)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis

@@

@@

  train.data <- left_join(trainsamples, traindata) %>%
    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex)

@@

  # LDA analysis
  # Fit the model
  model <- lda(treatment~ ., data = train.transformed)
  # Make predictions
  predictions <- model %>% predict(test.transformed)

  # Model accuracy
  print("model accuracy")
  print("predictions$class==test.transformed$treatment)")
  print(mean(predictions$class==test.transformed$treatment))

@@

```{r LDAplots}
library(kableExtra)

LDA.hyp$treatment <- factor(LDA.hyp$treatment, levels = alllevels)
LDA.hyp$predictedstage <- factor(LDA.hyp$predictedstage, levels = alllevels)

LDA.pit$treatment <- factor(LDA.pit$treatment, levels = alllevels)
LDA.pit$predictedstage <- factor(LDA.pit$predictedstage, levels = alllevels)

LDA.gon$treatment <- factor(LDA.gon$treatment, levels = alllevels)
LDA.gon$predictedstage <- factor(LDA.gon$predictedstage, levels = alllevels)

#myshapes = c("hypothalamus" = 20,  "pituitary" = 17,  "gonads" = 15)

a <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip) 
b <- ggplot(LDA.hyp, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 20) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

c <- ggplot(LDA.pit, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
d <- ggplot(LDA.pit, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 17) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

e <- ggplot(LDA.gon, aes(x = LD1, LD2, color = predictedstage)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)
f <- ggplot(LDA.gon, aes(x = LD1, LD2, color = treatment)) + geom_point(shape = 15) + theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)

plot_grid(a,b,c,d,e,f, nrow = 3)

g <- ggplot(LDA.hyp, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
  geom_point( ) + theme_B3() +
  theme(legend.position = "none") + 
  scale_color_manual(values = colorscharmaip) + 
  guides(color = guide_legend(show = FALSE))  +
  labs(subtitle = "hypothalamus")

h <- ggplot(LDA.pit, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
  geom_point( ) + theme_B3() +
  theme(legend.position = "none") + scale_color_manual(values = colorscharmaip)  + 
  guides(color = guide_legend(show = FALSE)) +
  labs(subtitle = "piuitary")

i <- ggplot(LDA.gon, aes(x = LD1, LD2, shape = predictedstage, color = treatment)) + 
  geom_point( ) + theme_B3() +
  theme(legend.position = "right") + scale_color_manual(values = colorscharmaip)  + 
  #guides(color = guide_legend(show = FALSE)) +
  guides(shape = guide_legend(ncol = 3),
         color = guide_legend(ncol = 3)) +
  labs(subtitle = "gonads")

gh <- plot_grid(g,h, nrow = 1)
ghi <- plot_grid(gh, i, nrow = 2)
ghi

plotfacetresults <- function(df, whichgroups, mysubstitle, myshape){

  p <- df %>%
    filter(treatment %in% whichgroups) %>%
    ggplot(aes(x = LD1, LD2,  color = predictedstage)) + 
  geom_point(shape = myshape ) + theme_B3() +
    stat_ellipse() +
  theme(legend.position = "none") + 
  scale_color_manual(values = colorscharmaip) + 
  guides(color = guide_legend(show = FALSE))  +
  labs(y = mysubstitle) + 
  facet_wrap(~treatment, nrow = 1)
  return(p)
}

j <- plotfacetresults(LDA.gon, levelstiming, "hypothalamus\n LD2", 20) + labs(x = NULL)
k <- plotfacetresults(LDA.pit, levelstiming, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
l <- plotfacetresults(LDA.gon, levelstiming, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow  = 1))

plot_grid(j,k,l, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")

m <- plotfacetresults(LDA.gon, levelsremoval, "hypothalamus\n LD2", 20) + labs(x = NULL)
n <- plotfacetresults(LDA.pit, levelsremoval, "pituitary\n LD2", 17) + theme(strip.text = element_blank()) + labs(x = NULL)
o <- plotfacetresults(LDA.gon, levelsremoval, "gonads\n LD2", 15) + theme(strip.text = element_blank()) + 
  theme(legend.position = "bottom") +
  guides(color = guide_legend(nrow  = 1))

plot_grid(m,n,o, nrow = 3, rel_heights = c(1.2,1,1.6), align = "v")

plotfacetresults2 <- function(df, whichgroups, mysubstitle, myshape){
  p <- df %>%
    filter(treatment %in% whichgroups) %>%
  ggplot(aes(x = treatment, fill = predictedstage)) +
  geom_bar(position = position_fill(reverse = F))  +
  facet_wrap(~sex) +
    theme_B3() +
  theme(legend.position = "none") +
  scale_fill_manual(values = colorscharmaip) +
  labs(y = mysubstitle) 
  return(p)
}

j <- plotfacetresults2(LDA.gon, levelstiming, "hypothalamus", 20) + labs(x = NULL) + 
  theme(axis.text.x = element_blank()) +
  theme(legend.position = "top") + guides(fill = guide_legend(nrow  = 1)) 
k <- plotfacetresults2(LDA.pit, levelstiming, "pituitary", 17) + 
  theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
l <- plotfacetresults2(LDA.gon, levelstiming, "gonads", 15) + theme(strip.text = element_blank()) + 
   labs(x = NULL)

plot_grid(j,k,l, nrow = 3, rel_heights = c(1.6,1,1.2))

m <- plotfacetresults2(LDA.gon, levelsremoval, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
    theme(legend.position = "top")  +  guides(fill = guide_legend(nrow  = 1))
n <- plotfacetresults2(LDA.pit, levelsremoval, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
o <- plotfacetresults2(LDA.gon, levelsremoval, "gonads", 15) +
  theme(strip.text = element_blank())  + labs(x = NULL)

plot_grid(m,n,o, nrow = 3, rel_heights = c(1.6,1,1.2))

m <- plotfacetresults2(LDA.gon, alllevels, "hypothalamus", 20) + labs(x = NULL) + theme(axis.text.x = element_blank()) +
    theme(legend.position = "top")  +  guides(fill = guide_legend(nrow  = 1))
n <- plotfacetresults2(LDA.pit, alllevels, "pituitary", 17) + theme(strip.text = element_blank()) + labs(x = NULL) + theme(axis.text.x = element_blank())
o <- plotfacetresults2(LDA.gon, alllevels, "gonads", 15) +
  theme(strip.text = element_blank())  + labs(x = NULL) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot_grid(m,n,o, nrow = 3, rel_heights = c(1.6,1,1.2))

>>>>

---
title: "04_LDA_v2"
output: md_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(cowplot)
library(forcats)
library(caret) # LDA analysis
library(MASS) # LDA analysis
library(kableExtra)

@@

# add new columns

colData <- colData %>%
  mutate(newgrouping = fct_collapse(treatment,
         "control" = c("control"),
         "bldg" = c("bldg"),
         "nearhatch" = c("inc.d17", "hatch"),
         "earlyinc" = c("lay","inc.d3", "inc.d9"),
         "chickcare" = c("n5", "n9"),
         "manip" = maniplevels1))

@@

  train.data <- left_join(trainsamples, traindata) %>%
    dplyr::select(-V1, -bird, -tissue, -group, -study, -sex, -treatment)

@@

  # LDA analysis
  # Fit the model
  model <- lda(newgrouping~ ., data = train.transformed)
  # Make predictions
  predictions <- model %>% predict(test.transformed)

  # Model accuracy
  print("model accuracy")
  print("predictions$class==test.transformed$newgrouping)")
  print(mean(predictions$class==test.transformed$newgrouping))

@@

```{r LDAplotsNew}
library(kableExtra)

LDA.hyp$treatment <- factor(LDA.hyp$treatment, levels = alllevels)
LDA.hyp$predictedstage <- factor(LDA.hyp$predictedstage, levels = levelsnewgrouping)

LDA.pit$treatment <- factor(LDA.pit$treatment, levels = alllevels)
LDA.pit$predictedstage <- factor(LDA.pit$predictedstage, levels = levelsnewgrouping)

LDA.gon$treatment <- factor(LDA.gon$treatment, levels = alllevels)
LDA.gon$predictedstage <- factor(LDA.gon$predictedstage, levels = levelsnewgrouping)

plotpredictions <- function(df, mysubstitle){
  ggplot(df, aes(x = treatment, fill = predictedstage)) +
  geom_bar(position = position_fill(reverse = F))  +
  facet_wrap(~sex) +
    theme_B3() +
  theme(legend.position = "none",
        axis.text.x = element_blank()) +
  scale_fill_manual(values = colorsnewgrouping) +
  labs(y = mysubstitle) 
}

a <- plotpredictions(LDA.hyp, "hypothalamus") + theme(legend.position = "top") 
b <- plotpredictions(LDA.pit, "pituitary") + theme(strip.text = element_blank())
c <- plotpredictions(LDA.gon, "gonads") + theme(axis.text.x = element_text(angle = 45, hjust = 1), strip.text = element_blank())

plot_grid(a,b,c, rel_heights = c(1.8,1,1.5), nrow = 3)
